---
title: "Static Vis"
author: "Daniel Krasnov"
date: "2023-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# breast cancer

https://www.kaggle.com/datasets/yasserh/breast-cancer-dataset

```{r}
library(tidyverse)
library(klaR)
library(MLmetrics)
bc <- read.csv("C:\\Users\\danie\\Desktop\\2023 W1\\DATA 419C\\Project\\DATA_419_Project\\breast-cancer.csv") 
```

```{r}
bc <- na.omit(bc)
bc <- bc[,-1]
bc$diagnosis <- factor(bc$diagnosis)
bc

as.data.frame(summary(bc))
```

id
Unique ID

diagnosis
Target: M - Malignant B - Benign

radius_mean
Radius of Lobes

texture_mean
Mean of Surface Texture

perimeter_mean
Outer Perimeter of Lobes

area_mean
Mean Area of Lobes

smoothness_mean
Mean of Smoothness Levels

compactness_mean
Mean of Compactness

concavity_mean
Mean of Concavity

concave points_mean
Mean of Cocave Points




```{r, message=FALSE, warning=FALSE}
# plot(bc[,2:11])
# library(GGally)
# library(ggplot2)
# # ggpairs(bc[,2:11])
# data <- bc[,1:11]
# ggpairs(data, columns = 1:11, aes(color = diagnosis, alpha = 0.5),
#         upper = list(continuous = wrap("cor", size = 2.5)))
```


```{r}
bc <- bc[,1:11]
rda_out <- rda(diagnosis~.,bc)
# Calculate and format results
post <- predict(rda_out)$posterior
class <- predict(rda_out)$class
y_pred <- as.numeric(predict(rda_out)$class)
y_true <- as.numeric(bc$diagnosis)

ConfusionMatrix(predict(rda_out)$class, bc$diagnosis)

res <-
data.frame(
"Precision" = Precision(y_true, y_pred),
"Recall" = Recall(y_true, y_pred),
"F1_Score" = F1_Score(y_true, y_pred)
)
res
```
## pca

```{r}
pca.out <- prcomp(bc[,-1],scale. = TRUE)
cum_prop_var <- summary(pca.out)$importance[3,2]
loadi <- round(pca.out$rotation[,1:2],4)
loadi
```

```{r}
library(klaR)
scores <- as.data.frame(pca.out$x)[,1:2]
df <- cbind(scores,bc$diagnosis)
colnames(df)[3] <- "diagnosis"
rda_out.pca <- rda(diagnosis~.,df)
rda_out.pca
```

```{r}
library(plotly)
# Set up grid
    x <- seq(-10, 4, length.out = 100)
    y <- seq(-8, 4, length.out = 100)
    grid <- expand.grid(x = x, y = y)
    colnames(grid) <- c("PC1", "PC2")
    
    # Get grid predictions
    prd = as.numeric(predict(rda_out.pca, newdata = grid)$class)
    
    
    # PCA preds
    scores <- as.data.frame(pca.out$x)[,1:2]
    df <- cbind(scores,bc$diagnosis)
    colnames(df)[3] <- "diagnosis"
    rda_out.pca <- rda(diagnosis~.,df)
    
    plot1 <-
      plot_ly()|> add_markers(
        data = scores,
        x = scores$PC1,
        y = scores$PC2,
        z = 0,
        opacity = 1,
        color = ~ bc$diagnosis,
        marker = list(size = 4),
        colors = "viridis",
        showlegend = FALSE
      ) |> add_markers(
        x = grid$PC1,
        y = grid$PC2,
        z = 0,
        color = ~prd,
        opacity=0.03,
        showlegend = FALSE
      )
    
plot1
```




```{r}
loadi <- round(pca.out$rotation[,1:2],4)
# loadi[abs(loadi)<0.001] <- NA
loadi
```



```{r}
scores <- as.data.frame(pca.out$x)[,1:2]
df <- cbind(scores,bc$diagnosis)
colnames(df)[3] <- "diagnosis"
rda(diagnosis~.,df)
```

## nmf

```{r}
#x: the non-negative data (if any negatives exist, the whole data is shifted)
#q: the number of factors/bases
#eps: convergence criteria (lack of progress on sum squared error)
#maxit: convergence criteria (max iteration when lack of progress not met)
#w: n by q scores on factors
#h: q by p observed factors/bases 
#By default (when w, h NULL) both w and h are initialized randomly
nmf <- function(x, q, eps=0.001, maxit=2000, w=NULL, h=NULL){
  n <- nrow(x)
  p <- ncol(x)
  if(any(x<0)){x <- as.matrix(x)+abs(min(x))}
  else{x <- as.matrix(x)}
  if(is.null(w)){
    w <- matrix(runif(n*q, min(x), max(x)), n, q)
  }
  if(is.null(h)){
    h <- matrix(runif(p*q, min(x), max(x)), q, p)
  }
  ed <- sum((x-w%*%h)^2)
  conv <- FALSE
  ctr <- 1
  while(!conv){
    ctr <- ctr+1
    h <- h * (t(w) %*% x) / (t(w) %*% w %*% h) 
    w <- w * (x %*% t(h)) / (w %*% h %*% t(h))
    wh <- w%*%h
    ed[ctr] <- sum((x-wh)^2)
    if((ed[ctr-1]-ed[ctr] < eps)|(ctr==maxit)){
      conv <- TRUE
    }
  }
  list(ed=ed, w=w, h=h, x=x)
}
```

```{r}
library(plotly)
    set.seed(87460945)
    # NMF preds
    bc_temp <- scale(bc[,-1], center = FALSE, scale = TRUE) 
    nmf.out <- nmf(bc_temp,2)
    scores <- as.data.frame(nmf.out$w)
    scores$diagnosis <- bc$diagnosis
    rda_out.nmf <- rda(diagnosis~.,scores)

    summary(scores)
    
    x <- seq(0, 8, length.out = 100)
    y <- seq(0, 4, length.out = 100)
    grid2 <- expand.grid(x = x, y = y)
    colnames(grid2) <- c("V1", "V2")
    
    # Get grid predictions
    prd = as.numeric(predict(rda_out.nmf, newdata = grid2)$class)
    
  
    
    plot1 <-
      plot_ly()|> add_markers(
        data = scores,
        x = scores$V1,
        y = scores$V2,
        z = 0,
        opacity = 1,
        color = ~ bc$diagnosis,
        marker = list(size = 4),
        colors = "viridis",
        showlegend = FALSE
      ) |> add_markers(
        x = grid2$V1,
        y = grid2$V2,
        z = 0,
        color = ~prd,
        opacity=0.03,
        showlegend = FALSE
      )

plot1
```


```{r}
library(mvtnorm)
# sigma1 <- matrix(c(1, 0, 0, 1), 2, 2, byrow = TRUE)
    sigma1 <- matrix(c(2, 0, 0, 3), 2, 2, byrow = TRUE)
    mu1 <- c(-1, -1)

    sigma2 <- matrix(c(1, 0, 0, 1), 2, 2, byrow = TRUE)
    mu2 <- c(0, 0)

    sigma3 <- matrix(c(0.5, 0, 0, 0.5), 2, 2, byrow = TRUE)
    mu3 <- c(1, 1)

    # DGP
    group1 <- data.frame(rmvnorm(100, mean = mu1, sigma = sigma1))
    group2 <- data.frame(rmvnorm(100, mean = mu2, sigma = sigma2))
    group3 <- data.frame(rmvnorm(100, mean = mu3, sigma = sigma3))
    groups <- c(rep("G1", 100), rep("G2", 100), rep("G3", 100))

    dgp <-
      data.frame(
        "G" = groups,
        "X1" = c(group1$X1, group2$X1,group3$X1),
        "X2" = c(group1$X2, group2$X2,group3$X2)
      )

    # Set up grid
    x <- seq(-5, 5, length.out = 100)
    y <- seq(-5, 5, length.out = 100)
    grid <- expand.grid(x = x, y = y)
    colnames(grid) <- c("X1", "X2")

    # Calculate densities values for dists
    pdf_values1 <- dmvnorm(grid, mean = mu1, sigma = sigma1)
    pdf_values2 <- dmvnorm(grid, mean = mu2, sigma = sigma2)
    pdf_values3 <- dmvnorm(grid, mean = mu3, sigma = sigma3)

    # Reshape the density values to match the grid
    density1 <- matrix(pdf_values1, nrow = length(x))
    density2 <- matrix(pdf_values2, nrow = length(x))
    density3 <- matrix(pdf_values3, nrow = length(x))

    # Fit RDA model
    rda.out <-
      rda(G ~ .,
          dgp,
          gamma = 1,
          lambda = 1)

    
    
    
    # Get grid predictions
    prd = as.numeric(predict(rda.out, newdata = grid)$class)
    
    # Get mixture model density
    unpooled_sigma <- rda.out$covariances
mean_mle <- rda.out$means
pooled_sigma <- rda.out$covpooled
gamma <- rda.out$regularization[1]
lambda <-  rda.out$regularization[2]

sigma_lambda_1 <- (1- lambda) * unpooled_sigma[,,1] + lambda * pooled_sigma
sigma_lambda_2 <- (1- lambda) * unpooled_sigma[,,2] + lambda * pooled_sigma
sigma_lambda_3 <- (1- lambda) * unpooled_sigma[,,3] + lambda * pooled_sigma

d <- 2
sigma_mixed_1 <- (1 - gamma) * sigma_lambda_1 + gamma * (1/d) * sum(diag((sigma_lambda_1))) *  diag(2)
sigma_mixed_2 <- (1 - gamma) * sigma_lambda_2 + gamma * (1/d) * sum(diag((sigma_lambda_2))) *  diag(2)
sigma_mixed_3 <- (1 - gamma) * sigma_lambda_3 + gamma * (1/d) * sum(diag((sigma_lambda_3))) *  diag(2)

# rda.out

# Assuming `data` is your simulated dataset
pdf_group1 <- dmvnorm(grid, mean = mean_mle[,1], sigma = sigma_mixed_1)
pdf_group2 <- dmvnorm(grid, mean = mean_mle[,2], sigma = sigma_mixed_2)
pdf_group3 <- dmvnorm(grid, mean = mean_mle[,3], sigma = sigma_mixed_3)

prior_group1 <- rda.out$prior[1] 
prior_group2 <- rda.out$prior[2]
prior_group3 <- rda.out$prior[3]

mixed_pdf <-  prior_group1*pdf_group1 + prior_group2*pdf_group2 + prior_group3*pdf_group3

mixed_density <- matrix(mixed_pdf,nrow=length(x))


opac_slid <- 0.5
    plot1 <-
      plot_ly() |>
      add_surface(
        x = x,
        y = y,
        z = density1,
        colors = "viridis",
        opacity = opac_slid,
        showlegend = FALSE
      ) |> add_surface(
        x = x,
        y = y,
        z = density2,
        colors = "viridis",
        opacity = opac_slid,
        showlegend = FALSE
      )|> add_surface(
        x = x,
        y = y,
        z = density3,
        colors = "viridis",
        opacity = opac_slid,
        showlegend = FALSE
      )|> add_surface(
        x = x,
        y = y,
        z = mixed_density,
        colors = "viridis",
        opacity = 1,
        showlegend = FALSE
      ) |> add_markers(
        data = dgp,
        x = dgp$X1,
        y = dgp$X2,
        z = 0,
        opacity = 1,
        color = ~ dgp$G,
        marker = list(size = 4),
        showlegend = FALSE
      ) |> add_markers(
        x = grid$X1,
        y = grid$X2,
        z = 0,
        color =  ~ prd,
        opacity = 0.01,
        showlegend = FALSE

      )
plot1
    



```

